---
title: "Coding Assignment 2"
author: "Team 4"
date: "Due: 2024-10-20 23:59"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
#Put any packages you need here
knitr::opts_chunk$set(echo = TRUE)
```

A Florida health insurance company wants to predict annual claims for individual clients. The company pulls a random sample of 100 customers. The owner wishes to charge an actuarially fair premium to ensure a normal rate of return. The owner collects all of their current customer’s health care expenses from the last year and compares them with what is known about each customer’s plan. 

The data on the 100 customers in the sample is as follows:

-	Charges: Total medical expenses for a particular insurance plan (in dollars)
-	Age: Age of the primary beneficiary
-	BMI: Primary beneficiary’s body mass index (kg/m2)
-	Female: Primary beneficiary’s birth sex (0 = Male, 1 = Female)
-	Children: Number of children covered by health insurance plan (includes other dependents as well)
-	Smoker: Indicator if primary beneficiary is a smoker (0 = non-smoker, 1 = smoker)
-	Cities: Dummy variables for each city with the default being Sanford

Answer the following questions using complete sentences and attach all output, plots, etc. within this report.


```{r dataset, include=FALSE}
insurance_data <- read.csv("~/Documents/GitHub/ECO6416_Group04/Insurance_Data_Group4.csv", header = TRUE, sep = ",", quote = "\"", dec = ".", fill = TRUE)
```


## Question 1

Generate this regression equation:

$$Charges = \beta_{0}+\beta_{1}*Age+\beta_{2}*BMI+\beta_{3}*Children + \beta_{4}*Female + \beta_{5}*Smoker + \beta_{6}*WinterSprings + \beta_{7}*WinterPark + \beta_{8}*Oviedo$$

```{r q1}
model <- lm(Charges ~ Age + BMI + Children + Female + Smoker + WinterSprings + WinterPark + Oviedo, data = insurance_data)
summary(model)
```

Which independent variables test significant in a two-tailed test? Use $\alpha = 0.2$

The variables that test significant in a two-tailed test using $\alpha = 0.2$ are Age, BMI, Smoker, WinterSprings, WinterPark, and Oviedo. Their p-values are all less than 0.2, allowing us to reject the null hypothesis for these variables.

## Question 2

Interpret the coefficients of all the independent variables that tested significant.

For age, each additional year older increases the charges by $286.53, holding all other factors constant. 
For BMI, each additional kg/m2 of BMI increases the charges by $239.91, holding all other factors constant.
If a customer is a smoker, the charges increase by $23828.07, holding all other factors constant.
If the customer lives in Winter Springs, the charges will increase by $2382.96, holding all other factors constant.
If the customer lives in Winter Park, the charges will increase by $2189.25, holding all other factors constant. 
If the customer lives in Oviedo, the charges will increase by $2054.11, holding all other factors constant.

## Question 3

The insurance agent returns with another client:

- Age: 35
- BMI: 26
- Children: 1
- Gender: Female
- Smoker: No
- City: Sanford

Using the regression equation above, predict the amount of medical expenses associated with this policy. (Provide a 95% confidence interval as well)

```{r q3}
new_data <- data.frame(Age = 35, 
                       BMI = 26,
                       Children = 1,
                       Female = 1,
                       Smoker = 0,
                       WinterSprings = 0,
                       WinterPark = 0,
                       Oviedo = 0)
prediction <- predict(model, newdata = new_data, interval = "confidence", level = 0.95)

print(prediction)
```
The predicted medical expenses associated with this policy are $2,946.20.  
There is a 95% probability that medical expenses for this customer will cost between $784.94 and $5107.46. 

## Question 4

Compare the model from question 1, to a model that does not include the group of city dummies. Does including the cities reduce the SSE by a statistically significant amount? (check the partial F-test for significance of the dummies as a group). Use $\alpha = 0.05$

```{r q4}
reduced_insurance <- insurance_data[,-(7:9)]
model_reduced <- lm(Charges ~., data = reduced_insurance)
summary(model_reduced)

anova(model_reduced, model)

```

Including the cities does not reduce the SSE by a statistically significant amount. The partial F-test showed a p-value of 0.2274, which did not test statistically significant against $/alpha=0.05$. So, we fail to reject the null hypothesis. In order for the cities to test statistically significant, their p-values would have to decrease significantly to get below 0.05.
